---
output: 
  html_document: 
    keep_md: yes
---
# Business Intelligence Lab Submission Markdown

9/24/2023

-   Ngumi Joshua 136346
-   Aicha Mbongo 134141
-   Joseph Watunu 127559
-   Muema Ian 135863
-   Hakeem Alavi 134775

# BUSINESS INTELLIGENCE 2

## ExploratoryDataAnalysis

# Student Details {#student-details}

|                                                   |                                                              |
|---------------------------------|---------------------------------------|
| **Student ID Numbers and Names of Group Members** | 136346,127559,135863,134775,134141                           |
| **BBIT 4.2 Group**                                | Group C                                                      |
| **Course Code**                                   | BBT4206                                                      |
| **Course Name**                                   | Business Intelligence II                                     |
| **Program**                                       | Bachelor of Business Information Technology                  |
| **Semester Duration**                             | 21<sup>st</sup> August 2023 to 28<sup>th</sup> November 2023 |

# Setup Chunk {#setup-chunk}

**Note:** the following KnitR options have been set as the global defaults:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE,                        collapse = FALSE, tidy = TRUE)`.

> output:
>
> github_document:\
> toc: yes\
> toc_depth: 4\
> fig_width: 6\
> fig_height: 4\
> df_print: default
>
> editor_options:\
> chunk_output_type: console

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

# Loading of dataset

The X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset dataset contains 101 observations and 100 variables.Variables include gender, class.group, YOB, internet, and day.out e.t.c

```{r 1st}
library(readr)
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset <- read_csv("data/20230412-20230719-BI1-BBIT4-1-StudentPerformanceDataset.csv")
View(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)

```

## Dimensions

```{r 2nd}
dim(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)
```

## Data Types

Knowing the data types will help you to identify the most appropriate,visualization types and algorithms that can be applied. It can also help you to identify the need to convert from categorical data (factors) to integers or vice versa where necessary

```{r 3rd }

sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset, class)
```

# Descriptive Statistics

Descriptive statistics are a set of numerical and graphical techniques used to summarize and present essential characteristics of a dataset. These statistics include measures like mean, median, and standard deviation.Some of the common measures include count,percentile, frequency e.t.c

## Measures of Frequency

Measures of frequency in statistics provide insights into how often different values or categories occur within a dataset.These measures help analysts understand the distribution and prevalence of data points, aiding in statistical analysis and decision-making.

```{r 4th }
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq <- X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$class_group
cbind(frequency = table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq),
      percentage = prop.table(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq)) *100)

X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq <- X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$alcohol_or_narcotics
cbind(frequency = table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq),
      percentage = prop.table(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq)) *100)

X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq <- X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$financial_wellness
cbind(frequency = table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq),
      percentage = prop.table(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq)) *100)

X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq <- X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$romantic_relationships
cbind(frequency = table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq),
      percentage = prop.table(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq)) *100)

X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq <- X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$gender
cbind(frequency = table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq),
      percentage = prop.table(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq)) *100)


```

# Measures of central tendency

The following code reveals the most frequently occurring performance level in the dataset. Additionally, it identifies the performance level of students who dropped the Business program based on the dataset.

```{r 12th}
student_performance_dataset_mode <- names(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$gender))[
  which(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$gender) == max(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$gender)))
]
print(student_performance_dataset_mode)

student_performance_dataset_mode <- names(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$drop_bi_now))[
  which(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$drop_bi_now) == max(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$drop_bi_now)))
]
print(student_performance_dataset_mode)

```

## Measures of Distribution/Dispersion/Spread/Scatter/Variability

Measures of distribution, dispersion, or variability quantify how data points in a dataset are spread out from the central tendency.They help assess the degree of data spread, aiding in understanding data variability and making informed decisions. Common measures include range, variance, standard deviation, and interquartile range.

```{r 5th}
summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)

```

### Measure the standard deviation of each variable

This code computes the standard deviation for each variable in the dataset.

```{r 6th }
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset[, c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99)], sd)

```

### The variance of each variable

This code computes the standard deviation for each variable in the dataset.

```{r 7th code}
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset[, c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99)], var)

```

### Measure the kurtosis of each variable

The Kurtosis informs you of how often outliers occur in the results.

```{r 8th}
if (!is.element("e1071", installed.packages()[, 1])) {
  install.packages("e1071", dependencies = TRUE)
}
require("e1071")
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset [, c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99)],kurtosis, type=2)

```

### Measure the skewness of each variable

The skewness informs you of the asymmetry of the distribution of results.Skewness between -0.4 and 0.4 (inclusive) implies that there is no skew.In the distribution of results; the distribution of results is symmetrical.It is a normal distribution.Skewness above 0.4 implies a positive skew; a right-skewed distribution.Skewness below -0.4 implies a negative skew; a left-skewed distribution.

```{r 9th}
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset [c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99)],skewness,type=2)

```

# Measures of Relationship

### Measures of covariance

Measures of covariance assess how two variables change together. Positive values indicate a direct relationship, negative values an inverse one.

```{r 10th}
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_cov <- cov(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset [c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99)])
View(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_cov)
```

### Measure the correlation between variables

Covariance measures how two variables move in relation to each other. Positive values indicate they tend to increase together, while negative values suggest one goes up as the other goes down.

```{r 11th}
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_cor <- cor(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset [c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99)])
View(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_cor)
```
